Discriminating the Influence of Correlated Factors \\from Multivariate Observations: the Back-to-Back Regression

### 2020-05-20 paper/response.tex

\documentclass[preprint,12pt,3p]{elsarticle}

\usepackage{lineno,hyperref}
% \modulolinenumbers[5]

\usepackage{amsmath, amsfonts, microtype, xcolor, tikz, graphicx, amsthm}
\usepackage[ruled, linesnumbered]{algorithm2e}
\usepackage{times,wrapfig}
\usepackage{cleveref}

\begin{document}
\title{Response to reviewers}

\section{Reviewer 1}
\begin{quote}\textbf{
    There is one major limitation for the broader application of this approach and
    that's demonstrating its effectiveness in fMRI data. The majority of
    neuroimaging studies are done using fMRI (and I do not say this to discount the
    other valuable tools of EEG, MEG, NIRS, etc.). To get other neuroscientists to
    grab onto this approach, I recommend the authors perform another analysis using
    fMRI data. My suggestion for the authors would be to use data from the Human
    Connectome Project (http://www.humanconnectomeproject.org/). The outcome
    variable used with this approach is not important as so much the demonstration
    that their technique is effective for fMRI data. Without including fMRI data,
    the approach comes off as useful but speculative for fMRI data, which
    drastically limits its impact potential.
}
\end{quote}

We thank Reviewer 1 for this remark. We now apply the same analyses to
fMRI recordings of 100 subjects performing a reading experiment similar to the
MEG study described in original manuscript. We kept the exact same methods and parameters, and
used the fmriprep package to perform the preprocessing.

@@

Overall, the results consistently show that B2B compares favorably to other methods.
Furthermore, these novel analyses complements the temporal results obtained with
MEG. Specifically, the effects of Word Length are predominantly observed over the early visual
cortices, and the Word Frequency are predominantly observed over the fronto-temporal areas.

\section{Reviewer 2}
\begin{quote}\textbf{
    First, the authors justify their approach of using backward modeling by the
    ability to estimate common noise sources. However, this would also be possible
    using multivariate forward modeling. For that reason, I'm wondering how
    multivariate forward modeling would fare under these circumstances. I would find
    it highly informative to include a (similarly cross-validated) multivariate GLM
    to the analyses and results.
}
\end{quote}

@@

We thank Reviewer 2 for this remark. The multivariate GLM / MANOVA is indeed
a relevant analysis. We now indicate in supplementary materials the following analyses:

\begin{quote}
    "Following the recommendations of one of our reviewers, we implemented a
multivariate variant of the forward model, i.e. a MANOVA, using the statsmodels
implementation. MANOVA is primarily used as a
inferential statistics, and does not trivially convert to a predicting method.
Consequently, we did not find a way to compare MANOVA against B2B with the
$\Delta R$ evaluation. However, the effects of MANOVA are generally summarized
with the Wilk's Lambda stastistics or its transformation into an $F$-value.
For each searchlight, we thus use the $F$-values of the
Wilk's Lambda statistics as proxy for $\hat S$ and feeds it to a second-level
Wilcoxon signed-rank tests across subjects, like we did for the other models.

The results show that all factors, including the Dummy variable, are
systematically above chance level in all recorded brain regions (Supp Fig.). This result
can be explained by the large dimensionality of $Y$. Indeed, limiting the
searchlight to a 1mm radius did not lead to these spurious effects but provided
results similar to the Forward model.

Nonetheless, MANOVA does appear to capture some plausible effects. Indeed, the
$F$-values obtained for both Word Length and Word Frequency were weakly but
significantly higher than those obtained with the Dummy variable in the
occipital and temporal brain areas (Supp Fig.). This results
suggests that the effect size of the MANOVA can be biased and, thus, is not
valid for second-level statistics.

Overall, this suggests that MANOVA (1) can lead to positively biased estimates of
$\hat S$ (2) appears weaker than B2B in terms of second-level analysis across
subjects (3) misses the effect of Word Function detected with the Forward model,
and (4) does not trivially translate into a prediction tool. Together these
elements thus suggest that MANOVA is less suitable to the present objective than
B2B."

\end{quote}

\begin{quote}\textbf{
Likewise, while I applaud the authors efforts in conducting these analyses,
    I am wondering whether the impact of this method could not be made even
    stronger by adding more experimental findings. Are there any other openly
    available datasets that could be used to demonstrate the ability to
    disentangle known factors?
}
\end{quote}

@@

We agree with Reviewer 2. As indicated to Reviewer 1, we now successfully apply
the same analyses to the fMRI recordings of 100 subjects reading word sequences.

\begin{quote}\textbf{
Rather than using the feature importance delta R, is it possible to just
    report the parameter estimate E hat at each time point? What would be the
    downside of this?
}
\end{quote}

@@

It is possible to report $\hat E$ (now relabelled $\hat S$), however they are
interpretable only if the second regression $H$ is not regularized.
We now added a Supplementary Figure to that effect.

Note, however, that $\hat E$ statistics are not in the same unit as the forward $\hat H$
coefficients (a.k.a betas values in GLM). Consequently, it is not straightforward to
compare these two metrics directly. As a proxy, we thus report their respective
second-level p-values across subjects. The associated figure shows that B2B's $\hat S$
coefficients compare favorably to the Forward $\hat H$ coefficient and the MANOVA
coefficients.

\begin{quote}\textbf{
The authors describe their method as unbiased but then introduce two
    regularization parameters, which will bias the estimated coefficients towards
    zero. While the statement "B2B leads to unbiased (i.e. zeros-centered) scalar
    coefficients for non-causal features" is correct for the mean of those
    coefficients, the variance will also be biased. I think these are important
    criteria to consider.
}
\end{quote}

@@

We agree with this remark. We now clarify the meaning and implication of the
bias in the problem statement:

"Note that $\hat S$ is unbiased, in the sense that it is centered around zero when there is no
effect, only if the second regression $H$ is not regularized. Second-level
statistics testing whether $\hat S$ is superior to 0 are thus only valid if $H$
is not regularized."

\begin{quote}\textbf{
The results of backward regression are not shown in Figure 4, but it would be
    nice to see those results. Is there a possibility to add them to Figure 4 in a
    separate panel?
}
\end{quote}
We now added the decoding performance to a supplementary figure. The latter show
that approximately the same cortical network can be used to decode Word Length
and Word Frequency. It also shows that the Dummy Variable can be decoded from
various brain areas, and from the visual cortex in particular.

\begin{quote}\textbf{
I think the authors could better explain the dimensions of the data (dy
    referring to the number of measurement channels, and dx to the number of
    independent variables). Perhaps the authors could add this nomenclature to their
    graphical illustration? Likewise, for the regression of Y onto X, it is unclear
    if this is a mass-univariate regression or a multivariate regression (i.e.
    taking the error covariance between Ys into account or not). This might be
    obvious to them, but is not necessarily obvious to the reader.
}
\end{quote}
We amended the text in multiple places to clarify this issue.

\begin{quote}\textbf{
I find the choice of the letter E for one of the causal factors unfortunate
    since it is typically reserved for the true error component of a multivariate
    model. Likewise, the choice of N is unfortunate since the letter n has multiple
    meanings in the text: one particular instance of n, or the number of samples.
    This is a very minor comment though.
}
\end{quote}
We now changed $E$ into $S$ for Selection matrix. We also changed the number of
samples into $m$, and keep $N$ as the Noise matrix. Thank you for these suggestions

\begin{quote}\textbf{
Typo line 28: "were" should read "where"?
}
\end{quote}
Yes this is now corrected.

\begin{quote}\textbf{
Line 181: do the authors refer to the Wilcoxon signed-rank test?
}
\end{quote}
Yes, this is now corrected.

\end{document}

### 2020-08-03 paper/appendix.tex

For the Backward Model, feature importance cannot be assessed because there is
no prediction.

@@

The decision threshold between "causal" and "non-causal" elements is thus a fraction
$\mu$, whose proportion arbitrarily depends on the necessity to favor type I and
type II errors. In practice, we cannot use this procedure for our MEG study,
because signal-to-noise ratio is unknown.

@@

Following the recommendations of one of our reviewers, we also implemented a
multivariate variant of the forward model, i.e. a MANOVA, using the statsmodels
implementation \citep{seabold2010statsmodels}. MANOVA is primarily used as a
inferential statistics, and does not trivially convert to a predicting method.
Consequently, we did not find a way to compare MANOVA against B2B with the
$\Delta R$ evaluation. However, the effects of MANOVA comes with its own effect
size statistics: the Wilk's Lambda. For each searchlight, we thus use the $F-values$ of the
Wilk's Lambda statistics as proxy for $\hat S$ and feeds it to a second-level
Wilcoxon signed-rank tests across subjects, like we did for the other models.

The results show that all factors, including the Dummy variable, are
systematically above chance level in all recorded brain regions \ref{fig:fmri_supp}. This result
can be explained by the large dimensionality of $Y$. Indeed, limiting the
searchlight to a 1mm radius did not lead to these spurious effects but provided
results similar to the Forward model.

Nonetheless, MANOVA does appear to capture some plausible effects. Indeed, the
$F-values$ obtained for both Word Length and Word Frequency were weakly but
significantly higher than those obtained with the Dummy variable in the
occipital and temporal brain areas \ref{fig:fmri_supp}.

@@

    \label{fig:fmri_supp}
  \end{center}
  \caption{Top rows. Pearson $R$ correlation obtained for a Backward decoding model and
  a Forward encoding model trained with one factor at a time. These models can
  not take into account the factor covariance, and thus lead to spurious effects
  (e.g. visual cortex effect for the dummy variable). Bottom rows. Second-
  level p-values across subjects for a Forward, a B2B and a Manova trained with
  all factors. The B2B achieves better p-values, without leading to spurious effects for
  the Dummy variable. The Manova leads to biased estimates due to its inability
  to deal with overfitting. The last row shows where the Manova F-values of each factor differs
  significantly from the F-values of the Dummy variable.}
\end{figure}

@@

To test whether each of the methods robustly scales to an increasingly
large number of potential causes $X$, we enhanced the four ad-hoc features
(word length, word frequency, word function, dummy variable) with another
ten features. These additional features corresponds to the first dimensions
of word embedding as provided by Spacy \citep{spacy2}. The results shown
in Figure \ref{fig:embeddings}, show that the feature importance of ad-hoc
features as derived by B2B remain unchanged and are actually improved.

>>>>

For the Backward Model, feature importance cannot be assessed because there is
no prediction combining multiple factors.

@@

The decision threshold between "causal" and "non-causal" elements is thus a fraction
$\mu$, whose proportion arbitrarily depends on the necessity to favor type I and
type II errors. In practice, we cannot use this procedure for our fMRI and MEG experiments,
because signal-to-noise ratio is unknown.

@@

Following the recommendations of one of our reviewers, we implemented a
multivariate variant of the forward model, i.e. a MANOVA, using the statsmodels
implementation \citep{seabold2010statsmodels}. MANOVA is primarily used as a
inferential statistics, and does not trivially convert to a predicting method.
Consequently, we did not find a way to compare MANOVA against B2B with the
$\Delta R$ evaluation. However, the effects of MANOVA are generally summarized
with the Wilk's Lambda stastistics or its transformation into an $F$-value.
For each searchlight, we thus use the $F$-values of the
Wilk's Lambda statistics as proxy for $\hat S$ and feeds it to a second-level
Wilcoxon signed-rank tests across subjects, like we did for the other models.

The results show that all factors, including the Dummy variable, are
systematically above chance level in all recorded brain regions (Fig.~\ref{fig:fmri_supp}). This result
can be explained by the large dimensionality of $Y$. Indeed, limiting the
searchlight to a 1mm radius did not lead to these spurious effects but provided
results similar to the Forward model.

Nonetheless, MANOVA does appear to capture some plausible effects. Indeed, the
$F$-values obtained for both Word Length and Word Frequency were weakly but
significantly higher than those obtained with the Dummy variable in the
occipital and temporal brain areas (Fig.~\ref{fig:fmri_supp}). This results
suggests that the effect size of the MANOVA can be biased and, thus, is not
valid for second-level statistics.

@@

    \label{fig:fmri_supp}
  \end{center}
  \caption{Top two rows. Pearson $R$ correlation obtained for a Backward decoding model and
  a Forward encoding model trained with one factor at a time. These models can
  not take into account the factor covariance, and thus lead to spurious effects
  (e.g. visual cortex effect for the dummy variable). Bottom five rows. Second-
  level p-values across subjects for the coefficients of the Forward, a B2B and a Manova trained with
  all factors, as well as for the $\Delta R$ of the B2B, for comparison.
  B2B achieves better p-values, without leading to spurious effects for
  the Dummy variable. The Manova leads to biased estimates due to its inability
  to deal with overfitting. The last row shows where the Manova's $F$-values of each factor differs
  significantly from the $F$-values estimated for the Dummy variable.}
\end{figure}

@@

To test whether each of the methods robustly scales to an increasingly
large number of potential causes $X$, we enhanced the four ad-hoc features
(word length, word frequency, word function, dummy variable) with another
ten features. These additional features corresponds to the first dimensions
of word embedding as provided by Spacy \citep{spacy2}. The MEG results shown
in Fig.~\ref{fig:embeddings}, show that the feature importance of ad-hoc
features as derived by B2B remain unchanged and are actually improved.

### 2020-08-03 paper/fmri_exps.tex

Next, we apply our method to brain imaging data from the anonymized multimodal
neuroimaging ``Mother Of all Unification Studies'' (MOUS) dataset
\citep{schoffelen2019204}. The dataset contains functional Magnetic Resonance
Imaging (fMRI) and magneto-encephalography (MEG)
recordings of each 102 healthy Dutch adults who performed a reading
task in the scanner. Thirteen subjects were excluded from the analysis (12 MEG and 1 fMRI)
because of technical difficulties reading the files.
%
Subjects were exposed to a rapid serial visual presentation of Dutch words. The
word lists consisted of 120 sentences, and scrambled lists of the same words.
Each word was presented on the computer screen for 351ms on average (min: 300ms,
max: 1400ms). Successive words were separated by a blank screen for 300ms, and
successive sentences were separated by an empty screen for a few (3-4) seconds.

@@

Many internal operations of \emph{fMRIPrep} use \emph{Nilearn} 0.6.2
\citep[RRID:SCR\_001362]{nilearn}, mostly within the functional
processing workflow. For more details of the pipeline, see
\href{https://fmriprep.readthedocs.io/en/latest/workflows.html}{the
section corresponding to workflows in \emph{fMRIPrep}'s documentation}.

@@

The preprocessed volumetric fMRI data was linearly projected to its closest surface
using nilearn \texttt{vol\_to\_surf} function with a 3 mm radius. A surface searchlight
analysis was then implemented by concatenating the surface
fMRI data within a 8mm radius of each vertex. For each subject and vertex, we
thus build an observation matrix $Y \in \mathbb{R}^{m \times d_y}$ of $m\approx$ 1400 words
by $d_y\approx40$ vertices per searchlight sphere. Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

@@

We aim to identify the word features that cause a variation in brain responses. We consider four distinct but collinear features.
%
First, 'Word Length' refers to the total number of letters. Word Length is expected to primarily cause a variation in the early evoked MEG responses (i.e. from 100 ms after stimulus onset) elicited by the retinotopically-tuned visual cortices (e.g. \citep{pegado2014timing}.).
%
Second, 'Word Frequency' indexes how frequently each word appears in Dutch and
was derived with the the Zipf logarithmic scale of \citep{van2014subtlex}
provided by the WordFreq package \citep{speerwordfreq}. Word Frequency is
expected to primarily cause a variation in the late evoked MEG responses (i.e.
from 400 ms) in the left frontal, temporal and parietal cortices
cortices \citep{kutas2011thirty,mccandliss2003visual}.
%
Third, 'Word Function' indicates whether each word is a content word (i.e. a
noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition,
a conjunction, a determinant, a pronoun or a numeral), and was derived from
Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has
not been thoroughly investigated with fMRI and MEG. While its causal contribution to
reading processes in the brain thus remains unclear, this lexical feature can
nonetheless be expected to present similar brain patterns to word frequency.
%
Finally, to verify that B2B and other methods would not inadequately identify non-causal features, we added a dummy feature, constructed from a noisy combination of Word Length and Word Frequency:
$dummy = z(length) + z(frequency) + \mathcal{N}$, where $z$ normalizes features
and $\mathcal{N}$ is a random vector sampling Gaussian distribution (all terms
thus have a zero-mean and a unit-variance).

To account for the delay of blood oxygenation level dependent responses, the
features were convolved using the Glover hemodynamic response function
of nilearn's \texttt{compute\_regressor} function an oversampling of 16 and default parameters \citep{nilearn}.

This procedure yields an $X \in \mathbb{R}^{m \times d_x}$ matrix of $m\approx$
840 TR (Repetition Time: 2 sec.) by
$d_x=4$ features for each subject. Each of the columns of $X$ is normalized to
have a zero-mean and a unit variance.

@@

We compare B2B to four standard methods: Forward regression, Backward
regression, Regularized CCA and PLS, as implemented in scikit-learn
\citep{sklearn} and pyrcca \citep{bilenko2016pyrcca}, and optimized with nested cross-validation over twenty $l2$
regularization parameters logarithmically spaced between $10^{-4}$ and $10^4$
(for regression methods) or 1 to 4 canonical components (for cross-decomposition
methods).

We used the feature importance described in Algorithm \ref{algorithm:b2b_fi} to assess the extent to which each feature $X_i$ specifically improves the prediction of held-out $Y$ data, using a 5-fold cross-validation (with shuffled trials to homogeneize the distributions between the training and testing splits).

Each model was implemented for each subject and each time sample independently.
Pairwise comparison between models were performed using a two-sided Wilcoxon
signed-rank test across subjects (m=90) using the average $\Delta R$ across time.
Corresponding effect sizes are shown in Figure~\ref{fig:meg_results}, and
p-values are reported below.

@@

Supplementary Figure \ref{fig:fmri_supp} shows that the Backward model decodes
the dummy variable well above chance level. In addition, it decodes both Word
Length and Word Frequency across large sections of the cortex, even though
these features are known to primarily influence early visual cortex and
associative cortices respectively. These results illustrate that while Backward
modeling achieve high sensitivity, it is not valid to estimate the specific
causal contribution of collinear features.

On the contrary, the $H$
coefficients of the Forward model reveals the expected spatial specificity of
Word Length and Word Frequency, peaking the visual and temporal cortices
respectively. However such Forward coefficients may be variable across
subjects, and could thus underestimate the significance of each factor at each
vertex, as evaluated with a  signed-rank tests across subjects. In principle, such
inter-subject variability could be less impactful with
multivariate observations methods such as CCA and PLS \citep{bilenko2016pyrcca,
king2018encoding}, but they fail to provide a single and interpretable coefficient
for each feature at each vertex.

In contrast, the $\hat S$ of an unbiased B2B (i.e. a B2B where $H$ is not
regularized) reaches higher levels of significance at each vertex
than the Forward $H$ \ref{fig:fmri_supp}. (Note that a direct
comparison between the Forward coefficients $H$ and those of B2B ($\hat S$)
is not feasible, as these two do not have the same units.)

To fairly compare the Forward, CCA, PLS and B2B methods on a common and robust
evaluation metric, robust to coefficient biases, we thus compare them on their ability to improve $Y$
predictions of out-of-sample data when one factor is introduced into the model.
This analysis leads to a $\Delta R$ as described in Algorithm
\ref{algorithm:b2b_fi}). As expected, the Forward, CCA, PLS
and B2B method predicted that the Dummy Variable does not improve the $Y$
prediction. This confirms that these methods accurately rule out the known
non-causal factor.

The average $\Delta R$ across subjects is displayed for each model and each
feature in \ref{fig:fmri_delta_r}. Overall, B2B favorably compares to baseline models.
To quantify this assessment, we compare for each subject separately
the average $\Delta R$ across vertices obtained between B2B and each
baseline model. To limit the inclusion of uninformative brain regions in this
summary, we restrict the analysis to vertices which can be reliably accounted
for by the Forward model ($p<.001$, not corrected across vertices)
\ref{fig:fmri_delta_r}. The results show that B2B significantly outperforms all
baseline models on all factors (all $p<0.0001$) but the dummy variable
(min $p=0.0852$).

>>>>

Next, we apply our method to brain imaging data from the anonymized multimodal
neuroimaging ``Mother Of all Unification Studies'' (MOUS) study
\citep{schoffelen2019204}. The dataset contains functional Magnetic Resonance
Imaging (fMRI) and magneto-encephalography (MEG)
recordings of 102 healthy Dutch adults who performed a reading
task in the scanner. Ten subjects were excluded from the analysis (9/102 MEG and 1/102 fMRI)
because of technical difficulties reading the files.
%
Subjects were exposed to a rapid serial visual presentation of Dutch words. The
word lists consisted of 120 sentences, and scrambled lists of the same words.
Each word was presented on the computer screen for 351ms on average (min: 300ms,
max: 1400ms). Successive words were separated by a blank screen for 300ms, and
successive sentences were separated by an empty screen for a few (3-4) seconds.

@@

Many internal operations of \emph{fMRIPrep} use \emph{Nilearn} 0.6.2
\citep[RRID:SCR\_001362]{Nilearn}, mostly within the functional
processing workflow. For more details of the pipeline, see
\href{https://fmriprep.readthedocs.io/en/latest/workflows.html}{the
section corresponding to workflows in \emph{fMRIPrep}'s documentation}.

@@

The preprocessed volumetric fMRI data was linearly projected to its closest surface
using Nilearn \texttt{vol\_to\_surf} function with a 3 mm radius and the
Freesurfer 'fsaverage5' surface. A surface searchlight
analysis was then implemented by concatenating the surface
fMRI data within a 8mm radius of each vertex. For each subject and vertex, we
thus build an observation matrix $Y \in \mathbb{R}^{m \times d_y}$ of $m\approx$ 1400 words
by $d_y\approx40$ vertices per searchlight sphere. Each of the columns of $Y$ is normalized to have zero mean and unit variance.

@@

We aim to identify the word features that cause a variation in brain responses. We consider four distinct but collinear features.
%
First, 'Word Length' refers to the total number of letters. Word Length is
expected to primarily cause a variation in the early evoked MEG responses
(i.e. from 100 ms after stimulus onset) primarily elicited by the
retinotopically-tuned visual cortices (e.g. \citep{pegado2014timing}.) but
expected across the full visual hierarchy (e.g. \citep{hagler2006spatial}).
%
Second, 'Word Frequency' indexes how frequently each word appears in Dutch and
here derives from the Zipf logarithmic scale of \citep{van2014subtlex}
provided by the WordFreq package \citep{speerwordfreq}. Word Frequency is
expected to primarily cause a variation in the late evoked MEG responses (i.e.
from 400 ms) in the left frontal, temporal and parietal
cortices \citep{kutas2011thirty,mccandliss2003visual}.
%
Third, 'Word Function' indicates whether each word is a content word (i.e. a
noun, a verb, an adjective or an adverb) or a function word (i.e. a preposition,
a conjunction, a determinant, a pronoun or a numeral), and here derives from
Spacy's part of speech tagger \citep{spacy2}. To our knowledge, this feature has
not been thoroughly investigated with fMRI and MEG. While its causal contribution to
reading processes in the brain thus remains unclear, this lexical feature can
nonetheless be expected to present similar brain patterns to word frequency.
%
Finally, to verify that B2B and other methods would not inadequately identify non-causal features, we added a dummy feature, constructed from a noisy combination of Word Length and Word Frequency:
$dummy = z(length) + z(frequency) + \mathcal{N}$, where $z$ normalizes features
and $\mathcal{N}$ is a random vector sampling Gaussian distribution (all terms
thus have a zero-mean and a unit-variance).

To account for the delay of blood oxygenation level dependent responses, these four
features were convolved using the Glover hemodynamic response function
of Nilearn's \texttt{compute\_regressor} function an oversampling of 16 and default parameters \citep{Nilearn}.

This procedure yields an $X \in \mathbb{R}^{m \times d_x}$ matrix of $m\approx$
840 TR (Repetition Time: 2 sec.) by $d_x=4$ factors for each subject.
Each of the columns of $X$ is normalized to have a zero-mean and a unit variance.

@@

We compare B2B to four standard methods: Forward regression, Backward
regression, Regularized CCA and PLS, as implemented in scikit-learn
\citep{sklearn} and Pyrcca \citep{bilenko2016Pyrcca}, and optimized with nested cross-validation over twenty $l2$
regularization parameters logarithmically spaced between $10^{-4}$ and $10^4$
(for regression methods) or 1 to 4 canonical components (for cross-decomposition
methods).

We used the feature importance described in Algorithm \ref{algorithm:b2b_fi} to assess the extent to which each feature $X_i$ specifically improves the prediction of held-out $Y$ data, using a 5-fold cross-validation (with shuffled trials to homogenize the distributions between the training and testing splits).

Each model was implemented for each subject and each time sample independently.
Pairwise comparison between models were performed using a two-sided Wilcoxon
signed-rank test across subjects using the average $\Delta R$ across time.
Corresponding effect sizes are shown in Fig.~\ref{fig:meg_results}, and
p-values are reported below.

@@

Supplementary Fig.~\ref{fig:fmri_supp} shows that the Backward model decodes
the dummy variable well above chance level. In addition, it decodes both Word
Length and Word Frequency across large sections of the cortex, even though
these features are known to primarily influence early visual cortex and
associative cortices respectively. In consequence, while Backward
modeling achieves high sensitivity, it is not valid to estimate the specific
causal contribution of collinear features.

On the contrary, the $H$
coefficients of the Forward model reveals the expected spatial specificity of
Word Length and Word Frequency, peaking in the visual and temporal cortices
respectively. However, such Forward coefficients may be variable across
subjects, and could thus underestimate the significance of each factor at each
vertex, as evaluated with a  signed-rank tests across subjects. In principle, such
inter-subject variability could be less impactful with
multivariate observations methods such as CCA and PLS \citep{bilenko2016Pyrcca,
king2018encoding}, but they fail to provide a single and interpretable coefficient
for each feature at each vertex.

In contrast, the $\hat S$ of an unbiased B2B (i.e. a B2B where $H$ is not
regularized) reaches higher levels of significance at each vertex
than the Forward $H$ (Supplementary Fig.~\ref{fig:fmri_supp}). (Note that a direct
comparison between the Forward coefficients $H$ and those of B2B ($\hat S$)
is challenging, because these two metrics do not have the same units.)

To fairly compare the Forward, CCA, PLS and B2B methods on a common and robust
evaluation metrics, robust to coefficient biases, we thus compare them on their ability to improve $Y$
predictions of out-of-sample data when one factor is introduced into the model.
This analysis leads to a $\Delta R$ as described in the synthetic experiment.
As expected, the Forward, CCA, PLS
and B2B method predicted that the Dummy Variable does not improve the $Y$
prediction. This confirms that these methods accurately rule out the known
non-causal factor.

The average $\Delta R$ across subjects is displayed for each model and each
feature in~\ref{fig:fmri_delta_r}. Overall, B2B favorably compares to baseline models.
To quantify this assessment, we compare for each subject separately
the average $\Delta R$ across vertices obtained between B2B and each
baseline model. To limit the inclusion of uninformative brain regions in this
summary, we restrict the analysis to vertices which can be reliably accounted
for by the Forward model ($p<.001$, not corrected across vertices)
The results, displayed in Fig.~\ref{fig:fmri_delta_r}, show that B2B
is significantly better than baseline models on all factors (all $p<0.0001$)
but the dummy variable (min $p=0.0852$).

### 2020-08-03 paper/meg_exps.tex

The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we
used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then
segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. The average responses across words is
displayed in Figure \ref{fig:meg_evoked}.
For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{m \times d_y}$ of $m\approx$ 2,700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

@@

We compare B2B against other methods following the same procedure as for the
fMRI experiments, except that the search light is applied across time samples.

@@

Figure \ref{fig:meg_results} shows, for each model, the effects obtained across
time (left) and subjects (right).

@@

To assess which model would be most sensitive to these causal discoveries, we
compared B2B to other models across subjects (Figure \ref{fig:meg_results}
right). For Word Length, B2B outperforms all models (all $p < 0.00001$) but CCA
($p=0.0678$). For Word Frequency, B2B outperforms all models (all $p < 0.0006$).
For "Word Function", B2B outperforms all models (all $p < 0.0015$). Overall,
these results show that B2B favorably compares against baseline models.

>>>>

One hundred and two subjects performed a similar reading task to the one
described above in the MEG scanner. The raw MEG data was bandpass-filtered between 0.1 and 40Hz using MNE-Python
default parameters \citep{gramfort2013meg, gramfort2014mne}. Specifically, we
used a zero-phase finite impulse
response filter (FIR) with a Hamming window and with transition bands of 0.1Hz
and 10Hz for the low and high cut-off frequencies. The raw data was then
segmented 100ms before word onset and 1s after
word onset ($t=0$ms corresponds to word onset). Finally, each resulting
segment was baseline-corrected between -100ms and 0ms, and decimated by 5 and
thus led a sampling frequency of 240Hz. The average responses across words is
displayed in Fig.~\ref{fig:meg_evoked}.
For each subject and each time sample relative to word onset, we
build an observation matrix $Y \in \mathbb{R}^{m \times d_y}$ of $m\approx$ 2,700 words
by $d_y=301$ MEG channels (273 magnetometers and 28 compensation channels). Each
of the columns of $Y$ is normalized to have zero mean and unit variance.

@@

We compare B2B against other methods following the same procedure as for the
fMRI experiments, except that the searchlight swipes across time samples, and
is trained across all MEG channels.

@@

Figure~\ref{fig:meg_results} shows, for each model, the effects obtained across
time (left) and subjects (right).

@@

We compare B2B to other models across subjects (Fig.~\ref{fig:meg_results}
right). For both Word Length and Word Frequency, B2B outperforms all models (all $p < 0.0001$).
For "Word Function", B2B outperforms all models (all $p < 0.0001$) but CCA ($p<0.0001$).
Overall, these results show that B2B generally compares favorably against baseline models.

### 2020-08-03 paper/paper.tex

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate independently and ii)
observations are multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a linear method designed to efficiently
estimate, from a set of correlated factors, those that most plausibly account for
multidimensional observations. First, we prove the consistency of B2B, its links to
other linear approaches, and show how it can provide a robust, unbiased and
interpretable scalar estimate for each factor.
Second, we use a variety of simulated data to show that B2B outperforms
forward modeling ("encoding"), backward modeling ("decoding") as well as
cross-decomposition modeling (i.e.. canonical
correlation analysis and partial least squares) on causal identification when
the factors and the observations are not orthogonal.
Finally, we apply B2B to a hundred magneto-encephalography recordings and to a hundred functional Magnetic
Resonance Imaging recordings acquired during a one hour reading task. B2B successfully disentangles
the respective contribution of collinear factors such as word length, word
frequency in the early visual and late associative cortical
responses respectively. These factors are better disentangled with B2B than with
other standard techniques. We discuss how the speed and generality of B2B
sets the foundations of an omnibus test for disentanglement of covarying
factors in multidimensional observations.
\end{abstract}

>>>>

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate independently and ii)
observations are multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a linear method designed to efficiently
estimate, from a set of correlated factors, those that most plausibly account for
multidimensional observations. First, we prove the consistency of B2B, its links to
other linear approaches, and show how it can provide a robust, unbiased and
interpretable scalar estimate for each factor.
Second, we use a variety of simulated data to show that B2B outperforms
forward modeling ("encoding"), backward modeling ("decoding") as well as
cross-decomposition modeling (i.e.. canonical
correlation analysis and partial least squares) on causal identification when
the factors and the observations are not orthogonal.
Finally, we apply B2B to a hundred magneto-encephalography recordings and to a hundred functional Magnetic
Resonance Imaging recordings acquired during a one hour reading task. B2B successfully disentangles
the respective contribution of collinear factors such as word length, word
frequency in the early visual and late associative cortical
responses respectively. These factors are better disentangled with B2B than with
other standard techniques. We discuss how the speed and generality of B2B
sets the foundations of an omnibus test to disentangle the causal contributions of covarying
factors in multidimensional observations.
\end{abstract}
